### 2025-07-25 paper/formalization.lagda.tex
agda code

\section{Agda Formalization}
\label{sec:formalization}

\begin{code}
data Type : Set where
  𝟘 𝟙 ⊥ ⊤ : Type
  ¡ ¿ : Type → Type
  _&_ _⊕_ _⊗_ _⅋_ : Type → Type → Type
\end{code}

\begin{code}
data Dual : Type → Type → Set where
  d-𝟘-⊤ : Dual 𝟘 ⊤
  d-⊤-𝟘 : Dual ⊤ 𝟘
  d-𝟙-⊥ : Dual 𝟙 ⊥
  d-⊥-𝟙 : Dual ⊥ 𝟙
  d-!-? : ∀{A B} → Dual A B → Dual (¡ A) (¿ B)
  d-?-! : ∀{A B} → Dual A B → Dual (¿ A) (¡ B)
  d-&-⊕ : ∀{A B A′ B′} → Dual A A′ → Dual B B′ → Dual (A & B) (A′ ⊕ B′)
  d-⊕-& : ∀{A B A′ B′} → Dual A A′ → Dual B B′ → Dual (A ⊕ B) (A′ & B′)
  d-⊗-⅋ : ∀{A B A′ B′} → Dual A A′ → Dual B B′ → Dual (A ⊗ B) (A′ ⅋ B′)
  d-⅋-⊗ : ∀{A B A′ B′} → Dual A A′ → Dual B B′ → Dual (A ⅋ B) (A′ ⊗ B′)
\end{code}

\begin{code}
dual-symm : ∀{A B} → Dual A B → Dual B A
\end{code}
\begin{code}[hide]
dual-symm d-𝟘-⊤ = d-⊤-𝟘
dual-symm d-⊤-𝟘 = d-𝟘-⊤
dual-symm d-𝟙-⊥ = d-⊥-𝟙
dual-symm d-⊥-𝟙 = d-𝟙-⊥
dual-symm (d-!-? p) = d-?-! (dual-symm p)
dual-symm (d-?-! p) = d-!-? (dual-symm p)
dual-symm (d-&-⊕ p q) = d-⊕-& (dual-symm p) (dual-symm q)
dual-symm (d-⊕-& p q) = d-&-⊕ (dual-symm p) (dual-symm q)
dual-symm (d-⊗-⅋ p q) = d-⅋-⊗ (dual-symm p) (dual-symm q)
dual-symm (d-⅋-⊗ p q) = d-⊗-⅋ (dual-symm p) (dual-symm q)
\end{code}