### 2025-07-27 paper/Reduction.lagda.tex
reduction

\begin{code}[hide]
open import Data.Bool using (Bool; true; false; if_then_else_)
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Relation.Binary.PropositionalEquality using (_≡_; refl; sym; cong; cong₂)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)
open import Data.List.Properties using (++-assoc)

open import Type
open import Context
open import Permutations
open import Process
open import Congruence
\end{code}

\subsection{Reduction}
\label{sec:reduction-agda}

\begin{code}[hide]
weakening : ∀{Γ Γ₁ Γ₂} (un : Un Γ₁) → Γ ≃ Γ₁ + Γ₂ → Process Γ₂ → Process Γ
weakening un p P = #process (+++# p) (aux un P)
  where
    aux : ∀{Γ₁ Γ₂} (un : Un Γ₁) → Process Γ₂ → Process (Γ₁ ++ Γ₂)
    aux un-[] P = P
    aux (un-∷ un) P = weaken (split-l +-unit-l) (aux un P)

contraction : ∀{Γ Γ₁ Γ₂} (un : Un Γ₁) → Γ ≃ Γ₁ + Γ₂ → Process (Γ₁ ++ Γ) → Process Γ
contraction un p P = #process (+++# p) (aux un (#process (#left (#sym (+++# p))) P))
  where
    aux : ∀{Γ₁ Γ₂} → Un Γ₁ → Process (Γ₁ ++ Γ₁ ++ Γ₂) → Process (Γ₁ ++ Γ₂)
    aux un-[] P = P
    aux {¿ A ∷ Γ₁} {Γ₂} (un-∷ un) P with contract (split-l +-unit-l) (#process (#shift {¿ A ∷ Γ₁} {¿ A} {Γ₁ ++ Γ₂}) P)
    ... | P₁ rewrite sym (++-assoc (¿ A ∷ Γ₁) Γ₁ Γ₂) with #process (#sym (#shift {Γ₁ ++ Γ₁} {¿ A})) P₁
    ... | P₂ rewrite ++-assoc Γ₁ Γ₁ (¿ A ∷ Γ₂) with aux un P₂
    ... | P₃ = #process #shift P₃
\end{code}

Just like structural precongruence, reduction is formalized as a binary relation
between processes that are well typed in the very same typing context. Thus, the
definition of reduction embeds the subject reduction result.

\begin{AgdaAlign}
\begin{code}
data _↝_ {Γ} : Process Γ → Process Γ → Set where
\end{code}

There is a constructor for each of the reduction rules in \Cref{tab:semantics}.
Let us comment a representative selection of them.

The constructor \AgdaInductiveConstructor{r-link} models \RLink.

\begin{code}
  r-link      : ∀{Δ A B} {P : Process (B ∷ Δ)}
                (d : Dual A B) (e : Dual A B) (p : Γ ≃ B , Δ) →
                cut d p (link e (split-l (split-r split-e))) P ↝ #process (#cons p) P
\end{code}

For generality, the rules quantifies over two proofs $d$ and $e$ of the same
duality relation between $A$ and $B$. Even though such proofs would necessarily
be the same, distinguishing them saves us a proof that they are equal whenever
we want to apply this reduction. 
%
Also note that the splitting
$(\AgdaInductiveConstructor{split-l}~(\AgdaInductiveConstructor{split-r}~\AgdaInductiveConstructor{split-e}))$
to which \AgdaInductiveConstructor{link} is applied is structured consistently
with \Cref{tab:semantics}, where the link $\Link\x\y$ being reduced is oriented
so that the restricted channel $x$ occurs on the left. 
%
Finally, the process $P\subst\y\x$ resulting from the reduction is subject to a
channel substitution in \Cref{tab:semantics}. In the Agda representation of
processes with nameless channels, the same effect is achieved by the permutation
$\AgdaFunction{\#cons}~p$, which ``renames'' (the position of) the free
occurrence of $x$ within $P$ with (the position of) $y$ encoded in the splitting
$p$.

The constructor \AgdaInductiveConstructor{r-close} models \RClose:

\begin{code}
  r-close     : ∀{P : Process Γ} (p₀ : Γ ≃ [] + Γ) (q₀ : Γ ≃ [] + Γ) →
                cut d-𝟙-⊥ p₀ close (wait (split-l q₀) P) ↝ P
\end{code}

While the \AgdaInductiveConstructor{close} constructor implicitly refers to the
only free channel occurring in a process of the form $\Close\x$, the
\AgdaInductiveConstructor{wait} constructor uses a splitting proof of the form
$(\AgdaInductiveConstructor{split-l}~q_0)$ to be sure that the referenced
channel is also the restricted one, and therefore matches the one of the
\AgdaInductiveConstructor{close} process.
%
Just like \AgdaInductiveConstructor{r-link} quantifies over multiple proofs for
the same duality relation, also \AgdaInductiveConstructor{r-close} (and several
other reduction constructors) quantifies over $p_0$ and $q_0$ which both prove
the splitting $\Context \simeq \AgdaInductiveConstructor{[]} + \Context$. Since
the left context is empty, these splittings must be equal and made of a sequence
of \AgdaInductiveConstructor{split-r} applications followed by
\AgdaInductiveConstructor{split-e}, but it is easier to quantify them separately
so that we do not have to worry about proving their equality whenever we wish to
apply this reduction.

The constructor \AgdaInductiveConstructor{r-select-l} models \RSelect when the
selected tag is $\InTag_1$, that is \AgdaInductiveConstructor{true} in Agda:

\begin{code}
  r-select-l  : ∀{Γ₁ Γ₂ A A′ B B′}
                {P : Process (A ∷ Γ₁)} {Q : Process (A′ ∷ Γ₂)} {R : Process (B′ ∷ Γ₂)}
                (d : Dual A A′) (e : Dual B B′)
                (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) →
                cut (d-⊕-& d e) p
                    (select true (split-l p₀) P)
                    (case (split-l q₀) Q R) ↝ cut d p P Q
\end{code}

There is not much more to note here (and in other omitted reduction rules),
except again for the multiple quantifications over the trivial splittings
$\Context_i \simeq \AgdaInductiveConstructor{[]} + \Context_i$ and the use of
\AgdaInductiveConstructor{split-l} to make sure that the channel referred to by
\AgdaInductiveConstructor{select} and \AgdaInductiveConstructor{case} is indeed
the one restricted by the cut.

\begin{code}[hide]
  r-select-r  :
    ∀{Γ₁ Γ₂ A A′ B B′}
    {P : Process (B ∷ Γ₁)} {Q : Process (A′ ∷ Γ₂)} {R : Process (B′ ∷ Γ₂)}
    (d : Dual A A′) (e : Dual B B′) (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) →
    cut (d-⊕-& d e) p
        (select false (split-l p₀) P)
        (case (split-l q₀) Q R) ↝ cut e p P R

  r-fork      :
    ∀{Γ₁ Γ₂ Γ₃ Δ A B A′ B′}
    {P : Process (A ∷ Γ₁)} {Q : Process (B ∷ Γ₂)} {R : Process (B′ ∷ A′ ∷ Γ₃)}
    (d : Dual A A′) (e : Dual B B′) (p : Γ ≃ Δ + Γ₃) (p₀ : Γ₃ ≃ [] + Γ₃)
    (q : Δ ≃ Γ₁ + Γ₂) (q₀ : Δ ≃ [] + Δ) →
    let _ , p′ , q′ = +-assoc-l p q in
    cut (d-⊗-⅋ d e) p
        (fork (split-l q₀) q P Q)
        (join (split-l p₀) R) ↝ cut d q′ P (cut e (split-r p′) Q R)

  r-client    :
    ∀{Γ₁ Γ₂ A A′}
    {P : Process (A ∷ Γ₁)} {Q : Process (A′ ∷ Γ₂)} (d : Dual A A′)
    (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
    cut (d-!-? d) p
      (server (split-l p₀) un P)
      (client (split-l q₀) Q) ↝ cut d p P Q

  r-weaken    :
    ∀{Γ₁ Γ₂ A A′}
    {P : Process (A ∷ Γ₁)} {Q : Process Γ₂}
    (d : Dual A A′) (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
    cut (d-!-? d) p
        (server (split-l p₀) un P)
        (weaken (split-l q₀) Q) ↝ weakening un p Q

  r-contract  :
    ∀{Γ₁ Γ₂ A A′}
    {P : Process (A ∷ Γ₁)} {Q : Process (¿ A′ ∷ ¿ A′ ∷ Γ₂)}
    (d : Dual A A′) (p : Γ ≃ Γ₁ + Γ₂) (p₀ : Γ₁ ≃ [] + Γ₁) (q₀ : Γ₂ ≃ [] + Γ₂) (un : Un Γ₁) →
    cut (d-!-? d) p
      (server (split-l p₀) un P)
      (contract (split-l q₀) Q) ↝
      contraction un p
        (cut (d-!-? d) ++≃+
             (server (split-l p₀) un P)
             (cut (d-!-? d) (split-r p) (server (split-l p₀) un P) Q))
\end{code}

Reduction is closed under cuts and by structural pre-congruence as per \RCut and
\RCong. The corresponding constructors that model these features are shown
below:

\begin{code}
  r-cut       : ∀{Γ₁ Γ₂ A B} {P Q : Process (A ∷ Γ₁)} {R : Process (B ∷ Γ₂)}
                (d : Dual A B) (q : Γ ≃ Γ₁ + Γ₂) → P ↝ Q →
                cut d q P R ↝ cut d q Q R

  r-cong      : ∀{P R Q : Process Γ} → P ⊒ R → R ↝ Q → P ↝ Q
\end{code}
\end{AgdaAlign}
In the following it is also useful to work with the reflexive, transitive
closure of $\red$, which is defined below:

\begin{code}
data _⇒_ {Γ} : Process Γ → Process Γ → Set where
  refl  : ∀{P : Process Γ} → P ⇒ P
  tran  : ∀{P Q R : Process Γ} → P ↝ Q → Q ⇒ R → P ⇒ R
\end{code}

\begin{code}[hide]
Reducible : ∀{Γ} → Process Γ → Set
Reducible P = ∃[ Q ] P ↝ Q
\end{code}