### 2025-11-27 paper/Type.lagda.tex
prose for types

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Nat
open import Data.Fin
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong; cong₂)
open import Agda.Builtin.Equality.Rewrite
\end{code}

\subsection{Type Representation}
\label{sec:type-agda}

The representation of types is fairly standard, with one constructor for each of
the forms described in \Cref{sec:calculus}. We use de Bruijn indices for
denoting type variables and we make sure that types are well-scoped by indexing
them with the number of variables that must be bound.

\begin{code}
data PreType : ℕ → Set where
  𝟘 𝟙 ⊥ ⊤          : ∀{n} → PreType n
  var rav          : ∀{n} → Fin n → PreType n
  ¡ ¿              : ∀{n} → PreType n → PreType n
  _&_ _⊕_ _⊗_ _⅋_  : ∀{n} → PreType n → PreType n → PreType n
  $∀ $∃            : ∀{n} → PreType (suc n) → PreType n
\end{code}

Note the constructors \AgdaInductiveConstructor{var} and
\AgdaInductiveConstructor{rav}, which respectively represent type variables and
their dual, and the quantifiers \AgdaInductiveConstructor{\$∀} and
\AgdaInductiveConstructor{\$∃} which increase the number of bound variables in
the scoped (pre)type.

The dual of a (pre)type is computed by the following \emph{function}
\AgdaFunction{dual}.

\begin{code}
dual : ∀{n} → PreType n → PreType n
dual 𝟘        = ⊤
dual 𝟙        = ⊥
dual ⊥        = 𝟙
dual ⊤        = 𝟘
dual (var x)  = rav x
dual (rav x)  = var x
dual (¡ A)    = ¿ (dual A)
dual (¿ A)    = ¡ (dual A)
dual (A & B)  = dual A ⊕ dual B
dual (A ⊕ B)  = dual A & dual B
dual (A ⊗ B)  = dual A ⅋ dual B
dual (A ⅋ B)  = dual A ⊗ dual B
dual ($∀ A)   = $∃ (dual A)
dual ($∃ A)   = $∀ (dual A)
\end{code}

It is straightforward to prove that duality is a an involution.

\begin{code}
dual-inv : ∀{n} {A : PreType n} → dual (dual A) ≡ A
\end{code}
\begin{code}[hide]
dual-inv {_} {𝟘} = refl
dual-inv {_} {𝟙} = refl
dual-inv {_} {⊥} = refl
dual-inv {_} {⊤} = refl
dual-inv {_} {var x} = refl
dual-inv {_} {rav x} = refl
dual-inv {_} {¡ A} = cong ¡ dual-inv
dual-inv {_} {¿ A} = cong ¿ dual-inv
dual-inv {_} {A & B} = cong₂ _&_ dual-inv dual-inv
dual-inv {_} {A ⊕ B} = cong₂ _⊕_ dual-inv dual-inv
dual-inv {_} {A ⊗ B} = cong₂ _⊗_ dual-inv dual-inv
dual-inv {_} {A ⅋ B} = cong₂ _⅋_ dual-inv dual-inv
dual-inv {_} {$∀ A} = cong $∀ dual-inv
dual-inv {_} {$∃ A} = cong $∃ dual-inv
\end{code}

In fact, this property is so important in the rest of the formalization that it
is worth defining an implicit rewriting rule so that Agda can autonomously apply
it whenever possible. This is achieved by means of the following pragma
directive.

\begin{code}
{-# REWRITE dual-inv #-}
\end{code}

\begin{code}[hide]
ext : ∀{m n} → (Fin m → Fin n) → Fin (suc m) → Fin (suc n)
ext ρ zero = zero
ext ρ (suc k) = suc (ρ k)

rename : ∀{m n} → (Fin m → Fin n) → PreType m → PreType n
rename ρ 𝟘 = 𝟘
rename ρ 𝟙 = 𝟙
rename ρ ⊥ = ⊥
rename ρ ⊤ = ⊤
rename ρ (var x) = var (ρ x)
rename ρ (rav x) = rav (ρ x)
rename ρ (¡ A) = ¡ (rename ρ A)
rename ρ (¿ A) = ¿ (rename ρ A)
rename ρ (A & B) = rename ρ A & rename ρ B
rename ρ (A ⊕ B) = rename ρ A ⊕ rename ρ B
rename ρ (A ⊗ B) = rename ρ A ⊗ rename ρ B
rename ρ (A ⅋ B) = rename ρ A ⅋ rename ρ B
rename ρ ($∀ A) = $∀ (rename (ext ρ) A)
rename ρ ($∃ A) = $∃ (rename (ext ρ) A)

exts : ∀{m n} → (Fin m → PreType n) → Fin (suc m) → PreType (suc n)
exts σ zero = var zero
exts σ (suc k) = rename suc (σ k)
\end{code}

Next we define the function \AgdaFunction{subst} that simultaneously substitutes
the type variables of a (pre)type with other (pre)types. Even though in practice
we will always substitute one variable at a time, it is technically easier to
substitutes \emph{all} variables of a (pre)type. The definition of
\AgdaFunction{subst} relies on some auxiliary functions for \emph{renaming} type
variables and \emph{lifting} substitutions across quantifiers. All such
definitions are straightforward adaptations of those describe by
\cite{KokkeWadler}.

\begin{code}
subst : ∀{m n} → (Fin m → PreType n) → PreType m → PreType n
\end{code}
\begin{code}[hide]
subst σ 𝟘 = 𝟘
subst σ 𝟙 = 𝟙
subst σ ⊥ = ⊥
subst σ ⊤ = ⊤
subst σ (var x) = σ x
subst σ (rav x) = dual (σ x)
subst σ (¡ A) = ¡ (subst σ A)
subst σ (¿ A) = ¿ (subst σ A)
subst σ (A & B) = subst σ A & subst σ B
subst σ (A ⊕ B) = subst σ A ⊕ subst σ B
subst σ (A ⊗ B) = subst σ A ⊗ subst σ B
subst σ (A ⅋ B) = subst σ A ⅋ subst σ B
subst σ ($∀ A) = $∀ (subst (exts σ) A)
subst σ ($∃ A) = $∃ (subst (exts σ) A)
\end{code}

Among all substitutions, we will use the one that substitutes the 0-indexed type
variable with a (pre)type. It is convenient to introduce this substitution once
and for all, which we do here.

\begin{code}
make-subst : ∀{n} → PreType n → Fin (suc n) → PreType n
make-subst A zero = A
make-subst A (suc k) = var k
\end{code}

Duality and substitutions are meant to commute.

\begin{code}
dual-subst : ∀{m n} {σ : Fin m → PreType n} {A : PreType m} →
             subst σ (dual A) ≡ dual (subst σ A)
\end{code}
\begin{code}[hide]
dual-subst {_} {_} {σ} {𝟘} = refl
dual-subst {_} {_} {σ} {𝟙} = refl
dual-subst {_} {_} {σ} {⊥} = refl
dual-subst {_} {_} {σ} {⊤} = refl
dual-subst {_} {_} {σ} {var x} = refl
dual-subst {_} {_} {σ} {rav x} = refl
dual-subst {_} {_} {σ} {¡ A} = cong ¿ (dual-subst {σ = σ} {A})
dual-subst {_} {_} {σ} {¿ A} = cong ¡ (dual-subst {σ = σ} {A})
dual-subst {_} {_} {σ} {A & B} = cong₂ _⊕_ (dual-subst {σ = σ} {A}) (dual-subst {σ = σ} {B})
dual-subst {_} {_} {σ} {A ⊕ B} = cong₂ _&_ (dual-subst {σ = σ} {A}) (dual-subst {σ = σ} {B})
dual-subst {_} {_} {σ} {A ⊗ B} = cong₂ _⅋_ (dual-subst {σ = σ} {A}) (dual-subst {σ = σ} {B})
dual-subst {_} {_} {σ} {A ⅋ B} = cong₂ _⊗_ (dual-subst {σ = σ} {A}) (dual-subst {σ = σ} {B})
dual-subst {_} {_} {σ} {$∀ A} = cong $∃ (dual-subst {σ = exts σ} {A})
dual-subst {_} {_} {σ} {$∃ A} = cong $∀ (dual-subst {σ = exts σ} {A})
\end{code}

Just like \AgdaFunction{dual-inv}, \AgdaFunction{dual-subst} too is key in the
formalization that follows. Therefore, we add it to the set of implicit
rewriting rules used by Agda so that we do not have to think about it again.

\begin{code}
{-# REWRITE dual-subst #-}
\end{code}

At this stage we are finally ready to define the type of closed pre-types, those
having no free type variables. We call \AgdaDatatype{Type} closed pre-types.
Hereafter, we will seldom use pre-types again.

\begin{code}
Type : Set
Type = PreType zero
\end{code}