### 2025-12-08 paper/agda.tex
restructuring presentation with new code

\section{Agda Formalization}
\label{sec:agda}

\input{Type}
\input{Context}
\input{Permutations}
\input{Process}
% \input{Congruence}
% \input{Reduction}
% \input{DeadlockFreedom}
% \input{Safety}