### 2025-12-11 paper/Context.lagda.tex
snapshot

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)
open import Relation.Unary
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong)

open import Type
\end{code}

\subsection{Context Representation}
\label{sec:context-agda}

We are going to adopt a nameless representation of channels whereby a channel is
identified by the position of its type in a typing context. This representation
is akin to using De Bruijn indices except that the index, instead of being
represented explicitly by a natural number, is computable from the proof that
the type of the channel belongs to the typing context.
%
Typing contexts are represented using lists of types, where the (polymorphic)
\AgdaDatatype{List} type and its constructors \AgdaInductiveConstructor{[]} and
\AgdaInductiveConstructor{\_∷\_} are defined in the module
\AgdaModule{Data.List} of Agda's standard library. We will keep using
$\ContextC$, $\ContextD$ and $\ContextE$ to range over typing contexts, even
though in the Agda formalization they are lists instead of finite maps.

\begin{code}
Context : Set
Context = List Type
\end{code}

The most important operation concerning typing contexts is \emph{splitting}. The
splitting of $\ContextC$ into $\ContextD$ and $\ContextE$, which we denote by
$\Splitting\ContextC\ContextD\ContextE$, represents the fact that $\ContextC$
contains all the types contained in $\ContextD$ and $\ContextE$, preserving both
their overall \emph{multiplicity} and also their relative \emph{order} within
$\ContextD$ and $\ContextE$. A \emph{proof} of
$\Splitting\ContextC\ContextD\ContextE$ shows how the types in $\ContextC$ are
distributed in $\ContextD$ and $\ContextE$ from left to right.

\begin{code}[hide]
infix  4 _≃_+_
infixr 8 _─∗_
infixr 9 _∗_
\end{code}
\begin{code}
data _≃_+_ : Context → Context → Context → Set where
  •   : [] ≃ [] + []
  <_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ A ∷ Δ + Θ
  >_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ Δ + A ∷ Θ
\end{code}

When splitting a context $\Context$ into $\ContextD + \ContextE$, for each type
in $\Context$ we use one of the prefix operators \AgdaInductiveConstructor{<}
and \AgdaInductiveConstructor{>} to indicate whether the type is meant to be
placed in $\ContextD$ or $\ContextE$.
%
For example, below is a proof of the splitting $\Splitting{[A, B, C,
D]}{[B]}{[A, C, D]}$.

\begin{code}[hide]
module _ where
  postulate A B C D : Type
\end{code}
\begin{code}
  splitting-example₁ : (A ∷ B ∷ C ∷ D ∷ []) ≃ (B ∷ []) + (A ∷ C ∷ D ∷ [])
  splitting-example₁ = > < > > •
\end{code}

Context splitting enjoys a number of expected properties. In particular, it is
easy to see that splitting is commutative and that the empty context/list is
both a left and right unit of splitting.

\begin{code}
+-comm    : ∀{Γ Δ Θ} → Γ ≃ Δ + Θ → Γ ≃ Θ + Δ
≫         : ∀{Γ} → Γ ≃ [] + Γ
≪         : ∀{Γ} → Γ ≃ Γ + []
\end{code}
\begin{code}[hide]
+-comm • = •
+-comm (< p) = > (+-comm p)
+-comm (> p) = < (+-comm p)

++≃+ : ∀{Γ Δ} → Γ ++ Δ ≃ Γ + Δ
++≃+ {[]}    {[]}    = •
++≃+ {[]}    {_ ∷ _} = > ++≃+
++≃+ {_ ∷ _} {_}     = < ++≃+

≫ = ++≃+ {[]}

≪ = +-comm ≫
\end{code}

Context splitting is also associative in a sense that is made precise below. If
we write $\ContextD + \ContextE$ for some $\ContextC$ such that $Γ ≃ Δ + Θ$, then
we can prove that $Γ_1 + (Γ_2 + Γ_3) = (Γ_1 + Γ_2) + Γ_3$.

\begin{code}
+-assoc-r  : ∀{Γ Δ Θ Δ′ Θ′} → Γ ≃ Δ + Θ → Θ ≃ Δ′ + Θ′ →
             ∃[ Γ′ ] Γ′ ≃ Δ + Δ′ × Γ ≃ Γ′ + Θ′
+-assoc-l  : ∀{Γ Δ Θ Δ′ Θ′} → Γ ≃ Δ + Θ → Δ ≃ Δ′ + Θ′ →
             ∃[ Γ′ ] Γ′ ≃ Θ′ + Θ × Γ ≃ Δ′ + Γ′
\end{code}
\begin{code}[hide]
+-assoc-r • • = [] , • , •
+-assoc-r (< p) q with +-assoc-r p q
... | _ , p′ , q′ = _ , < p′ , < q′
+-assoc-r (> p) (< q) with +-assoc-r p q
... | _ , p′ , q′ = _ , > p′ , < q′
+-assoc-r (> p) (> q) with +-assoc-r p q
... | _ , p′ , q′ = _ , p′ , > q′

+-assoc-l p q with +-assoc-r (+-comm p) (+-comm q)
... | Δ , r , p′ = Δ , +-comm r , +-comm p′
\end{code}

% It is often the case that the context $\ContextD$ in a splitting
% $\Splitting\ContextC\ContextD\ContextE$ is a singleton list $[A]$. We use this
% particular splitting to describe the presence of a channel of type $A$ in
% $\ContextC$ and giving a name $\ContextE$ to the residual context. For this
% particular case, we introduce the dedicated notation
% $\SimpleSplitting\ContextC{A}\ContextE$ which we find clearer.

% \begin{code}
% _∋_⊳_ : Context → Type → Context → Set
% Γ ∋ A ⊳ Δ = Γ ≃ [ A ] + Δ
% \end{code}
\begin{code}[hide]
+-empty-l : ∀{Γ Δ} → Γ ≃ [] + Δ → Γ ≡ Δ
+-empty-l • = refl
+-empty-l (> p) = cong (_ ∷_) (+-empty-l p)
\end{code}

Note that, when specifying a splitting $\Splitting\ContextC{[A]}\ContextE$, it
may be convenient to use \AgdaFunction{≫} as a shortcut for a sequence of
applications of \AgdaInductiveConstructor{>} after the $A$ type has been
reached. For instance, \AgdaFunction{splitting-example₁} can be written
equivalently and in a more compact way as shown below. More usages of
\AgdaFunction{≫} will be given in \Cref{sec:examples}.

\begin{code}[hide]
module _ where
\end{code}
\begin{code}
  splitting-example₂ : (A ∷ B ∷ C ∷ D ∷ []) ≃ [ B ] + (A ∷ C ∷ D ∷ [])
  splitting-example₂ = > < ≫
\end{code}

\begin{table}
  \begin{tabular}{|lll|}
    \hline
    \textbf{Notation} & \textbf{Definition} & \textbf{Meaning} \\
    \hline
    $\AgdaFunction{∀[}~P~\AgdaFunction]$ & $\forall\{x\} \to P~x$ & implicit universality \\
    $P~\AgdaFunction{⇒}~Q$ & $\lambda x \to P~x \to Q~x$ & implication \\
    $P~\AgdaFunction{∪}~Q$ & $\lambda x \to P~x~\AgdaDatatype{⊎}~Q~x$ & disjunction \\
    $P~\AgdaFunction{∩}~Q$ & $\lambda x \to P~x~\AgdaDatatype{×}~Q~x$ & conjunction \\
    $f~\AgdaFunction{⊢}~P$ & $\lambda x \to P~(f~x)$ & update \\
    $\AgdaFunction{U}$ & $\lambda x \to \AgdaModule{Data.Unit}.\AgdaDatatype{⊤}$ & universal
    set \\
    $\AgdaFunction{⋂[}~X~\AgdaFunction{:}~A~\AgdaFunction]~P$ & $\lambda x \to (X : A) \to P~X~x$ & infinitary conjunction \\
    \hline
  \end{tabular}
  \caption{\label{tab:predicates} Useful definitions from Agda's \AgdaModule{Relation.Unary} module.}
\end{table}

In the following we will make extensive use of predicates over contexts. For
this reason, it is worth recalling a number of definitions from the
\AgdaModule{Relation.Unary} module in Agda's standard library in
\Cref{tab:predicates}.

Context splitting is going to be pervasive in the subsequent development, to the
point that it contributes substantially to the complexity of the formalization.
For this reason, we invest some effort in defining a few abstractions that allow
us to occasionally hide context splittings and enhance code readability.
%
Inspired by separation logic~\cite{Hoare} and
following~\citet{RouvoetPoulsenKrebbersVisser20}, we define the \emph{separating
conjunction} $P \sand Q$ of two predicates $P$ and $Q$ over contexts:
\begin{code}
data _∗_ (P Q : Pred Context _) (Γ : Context) : Set where
  _⟨_⟩_ : ∀{Δ Θ} → P Δ → Γ ≃ Δ + Θ → Q Θ → (P ∗ Q) Γ
\end{code}

The predicate $P \sand Q$ holds for those contexts $\ContextC$ that can be split
into $\ContextD$ and $\ContextE$ where $P$ holds for $\ContextD$ and $Q$ holds
for $\ContextE$.
%
The constructor \AgdaInductiveConstructor{\_⟨\_⟩\_} has three explicit arguments
witnessing the splitting $\Splitting\ContextC\ContextD\ContextE$ along with
$P~\ContextD$ and $Q~\ContextE$.\Luca{Nei related fare presente che queste
definizioni seguono sostanzialmente \citet{RouvoetPoulsenKrebbersVisser20} ed
evidenziare le differenze.}

Along with \AgdaDatatype{∗} we define the \emph{separating implication} (also
known as ``magic wand'')
\begin{code}
_─∗_ : Pred Context _ → Pred Context _ → Context → Set
(P ─∗ Q) Δ = ∀{Θ Γ} → Γ ≃ Δ + Θ → P Θ → Q Γ
\end{code}
and prove that \AgdaFunction{─∗} can be used to curry \AgdaDatatype{∗}:
\begin{code}
curry∗ : ∀{P Q R} → ∀[ P ∗ Q ⇒ R ] → ∀[ P ⇒ Q ─∗ R ]
curry∗ F px σ qx = F (px ⟨ σ ⟩ qx)
\end{code}

To conclude the modeling of typing contexts, we define a predicate
\AgdaDatatype{Un} that holds for \emph{unrestricted} contexts, namely those
contexts solely made of types of the form $?A$. We will need this predicate in
the definition of a server, which must obey the typing rule \ServerRule of
\Cref{tab:typing-rules}.

\begin{code}
data Un : Context → Set where
  un-[]  : Un []
  un-∷   : ∀{A} → ∀[ Un ⇒ (`? A ∷_) ⊢ Un ]
\end{code}

It is easy to prove that $\Splitting\ContextC\ContextD\ContextE$ is unrestricted
if so are $\ContextD$ and $\ContextE$:

\begin{code}
∗-un : ∀[ Un ∗ Un ⇒ Un ]
\end{code}
\begin{code}[hide]
∗-un (un-[] ⟨ • ⟩ un-[])    = un-[]
∗-un (un-∷ un ⟨ < σ ⟩ un′)  = un-∷ (∗-un (un ⟨ σ ⟩ un′))
∗-un (un ⟨ > σ ⟩ un-∷ un′)  = un-∷ (∗-un (un ⟨ σ ⟩ un′))
\end{code}