### 2025-12-24 paper/Type.lagda.tex
revised types

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Nat
open import Data.Fin
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong; cong₂)
open import Agda.Builtin.Equality.Rewrite
\end{code}

\subsection{Type Representation}
\label{sec:type-agda}

The representation of types is standard, with one constructor for each of the
forms described in \Cref{sec:types}. We start by defining \emph{pre-types} as an
indexed data type $\AgdaDatatype{PreType}~n$ where $n$ is the number of bound
type variables and we use de Bruijn indices for denoting type variables. In this
way, we make sure that pre-types are well-scoped.

\begin{code}
data PreType : ℕ → Set where
  ⊤ 𝟘 ⊥ 𝟙          : ∀{n} → PreType n
  var rav          : ∀{n} → Fin n → PreType n
  _&_ _⊕_ _⅋_ _⊗_  : ∀{n} → PreType n → PreType n → PreType n
  `∀ `∃            : ∀{n} → PreType (suc n) → PreType n
  `! `?            : ∀{n} → PreType n → PreType n
\end{code}

Note the constructors \AgdaInductiveConstructor{var} and
\AgdaInductiveConstructor{rav}, which respectively represent type variables and
their dual, and the quantifiers \AgdaInductiveConstructor{`∀} and
\AgdaInductiveConstructor{`∃} which increase the number of bound type variables
in the scoped (pre)type.

The dual of a (pre)type is computed by the following function:

\begin{code}
dual : ∀{n} → PreType n → PreType n
dual ⊤        = 𝟘
dual 𝟘        = ⊤
dual ⊥        = 𝟙
dual 𝟙        = ⊥
dual (var x)  = rav x
dual (rav x)  = var x
dual (A & B)  = dual A ⊕ dual B
dual (A ⊕ B)  = dual A & dual B
dual (A ⅋ B)  = dual A ⊗ dual B
dual (A ⊗ B)  = dual A ⅋ dual B
dual (`∀ A)   = `∃ (dual A)
dual (`∃ A)   = `∀ (dual A)
dual (`! A)    = `? (dual A)
dual (`? A)    = `! (dual A)
\end{code}

It is straightforward to prove that duality is a an involution.

\begin{code}
dual-inv : ∀{n} {A : PreType n} → dual (dual A) ≡ A
\end{code}
\begin{code}[hide]
dual-inv {_} {⊤} = refl
dual-inv {_} {𝟘} = refl
dual-inv {_} {⊥} = refl
dual-inv {_} {𝟙} = refl
dual-inv {_} {var x} = refl
dual-inv {_} {rav x} = refl
dual-inv {_} {A & B} = cong₂ _&_ dual-inv dual-inv
dual-inv {_} {A ⊕ B} = cong₂ _⊕_ dual-inv dual-inv
dual-inv {_} {A ⅋ B} = cong₂ _⅋_ dual-inv dual-inv
dual-inv {_} {A ⊗ B} = cong₂ _⊗_ dual-inv dual-inv
dual-inv {_} {`∀ A} = cong `∀ dual-inv
dual-inv {_} {`∃ A} = cong `∃ dual-inv
dual-inv {_} {`! A} = cong `! dual-inv
dual-inv {_} {`? A} = cong `? dual-inv
\end{code}

This property is so important in the rest of the formalization that it is worth
defining an implicit rewriting rule so that Agda can autonomously apply it
whenever possible. This is achieved by means of the following pragma
directive.\footnote{The directive is effective provided that the option
\texttt{{-}{-}rewriting} is enabled, either globally when invoking Agda or in
the module's preamble.}

\begin{code}
{-# REWRITE dual-inv #-}
\end{code}

\begin{code}[hide]
ext : ∀{m n} → (Fin m → Fin n) → Fin (suc m) → Fin (suc n)
ext ρ zero = zero
ext ρ (suc k) = suc (ρ k)

rename : ∀{m n} → (Fin m → Fin n) → PreType m → PreType n
rename ρ ⊤ = ⊤
rename ρ 𝟘 = 𝟘
rename ρ ⊥ = ⊥
rename ρ 𝟙 = 𝟙
rename ρ (var x) = var (ρ x)
rename ρ (rav x) = rav (ρ x)
rename ρ (A & B) = rename ρ A & rename ρ B
rename ρ (A ⊕ B) = rename ρ A ⊕ rename ρ B
rename ρ (A ⅋ B) = rename ρ A ⅋ rename ρ B
rename ρ (A ⊗ B) = rename ρ A ⊗ rename ρ B
rename ρ (`∀ A) = `∀ (rename (ext ρ) A)
rename ρ (`∃ A) = `∃ (rename (ext ρ) A)
rename ρ (`! A) = `! (rename ρ A)
rename ρ (`? A) = `? (rename ρ A)

exts : ∀{m n} → (Fin m → PreType n) → Fin (suc m) → PreType (suc n)
exts σ zero = var zero
exts σ (suc k) = rename suc (σ k)
\end{code}

Next we define the function \AgdaFunction{subst} that simultaneously substitutes
the type variables of a (pre)type with other (pre)types. Even though in practice
we will always substitute one variable at a time, it is technically easier to
substitutes \emph{all} variables of a (pre)type. The definition of
\AgdaFunction{subst} relies on some auxiliary functions for \emph{renaming} type
variables and \emph{lifting} substitutions across quantifiers. All such
definitions are straightforward adaptations of those described by
\citet{KokkeSiekWadler20}.

\begin{code}
subst : ∀{m n} → (Fin m → PreType n) → PreType m → PreType n
\end{code}
\begin{code}[hide]
subst σ ⊤ = ⊤
subst σ 𝟘 = 𝟘
subst σ ⊥ = ⊥
subst σ 𝟙 = 𝟙
subst σ (var x) = σ x
subst σ (rav x) = dual (σ x)
subst σ (A & B) = subst σ A & subst σ B
subst σ (A ⊕ B) = subst σ A ⊕ subst σ B
subst σ (A ⅋ B) = subst σ A ⅋ subst σ B
subst σ (A ⊗ B) = subst σ A ⊗ subst σ B
subst σ (`∀ A) = `∀ (subst (exts σ) A)
subst σ (`∃ A) = `∃ (subst (exts σ) A)
subst σ (`! A) = `! (subst σ A)
subst σ (`? A) = `? (subst σ A)
\end{code}

Among all substitutions, we will use the one that substitutes the 0-indexed type
variable with a (pre)type. It is convenient to introduce this substitution once
and for all, which we do here.

\begin{code}
[_/] : ∀{n} → PreType n → Fin (suc n) → PreType n
[ A /] zero     = A
[ A /] (suc k)  = var k
\end{code}

Duality and substitutions are meant to commute.

\begin{code}
dual-subst : ∀{m n} {σ : Fin m → PreType n} {A : PreType m} →
             subst σ (dual A) ≡ dual (subst σ A)
\end{code}
\begin{code}[hide]
dual-subst {_} {_} {σ} {⊤} = refl
dual-subst {_} {_} {σ} {𝟘} = refl
dual-subst {_} {_} {σ} {⊥} = refl
dual-subst {_} {_} {σ} {𝟙} = refl
dual-subst {_} {_} {σ} {var x} = refl
\end{code}

It is worth looking at one case in the proof of \AgdaFunction{dual-subst},
namely when the type is a dualized type variable:

\begin{code}
dual-subst {_} {_} {σ} {rav x} = refl
\end{code}

Here we are supposed to prove
$\AgdaFunction{subst}~\sigma~(\AgdaFunction{dual}~(\AgdaInductiveConstructor{rav}~x))
  \equiv
  \AgdaFunction{dual}~(\AgdaFunction{subst}~\sigma~(\AgdaInductiveConstructor{rav}~x))$
  which is definitionally the same of $\sigma~x \equiv
  \AgdaFunction{dual}~(\AgdaFunction{dual}~(\sigma~x))$.
%
We could easily prove this equivalence by invoking \AgdaFunction{dual-inv}, but
thanks to the rewriting rule that we have added just above a use of
\AgdaInductiveConstructor{refl} suffices. In this case the saved effort is
negligible, but in later results, where it is necessary to use
\AgdaFunction{dual-inv} for rewriting part of the \emph{index} of some type
families, having an implicit rewriting rule allows us to avoid writing some
quite obscure Agda code.

\begin{code}[hide]
dual-subst {_} {_} {σ} {A & B} = cong₂ _⊕_ (dual-subst {σ = σ} {A}) (dual-subst {σ = σ} {B})
dual-subst {_} {_} {σ} {A ⊕ B} = cong₂ _&_ (dual-subst {σ = σ} {A}) (dual-subst {σ = σ} {B})
dual-subst {_} {_} {σ} {A ⅋ B} = cong₂ _⊗_ (dual-subst {σ = σ} {A}) (dual-subst {σ = σ} {B})
dual-subst {_} {_} {σ} {A ⊗ B} = cong₂ _⅋_ (dual-subst {σ = σ} {A}) (dual-subst {σ = σ} {B})
dual-subst {_} {_} {σ} {`∀ A} = cong `∃ (dual-subst {σ = exts σ} {A})
dual-subst {_} {_} {σ} {`∃ A} = cong `∀ (dual-subst {σ = exts σ} {A})
dual-subst {_} {_} {σ} {`! A} = cong `? (dual-subst {σ = σ} {A})
dual-subst {_} {_} {σ} {`? A} = cong `! (dual-subst {σ = σ} {A})
\end{code}

Just like \AgdaFunction{dual-inv}, \AgdaFunction{dual-subst} too is key in the
formalization that follows. Therefore, we add it to the set of implicit
rewriting rules used by Agda so that we do not have to think about this property
again.

\begin{code}
{-# REWRITE dual-subst #-}
\end{code}

At last we are ready to define the type of closed pre-types, those
having no free type variables. We call \AgdaDatatype{Type} closed pre-types.
From now on, we will seldom use pre-types again.

\begin{code}
Type : Set
Type = PreType zero
\end{code}