### 2025-12-24 paper/Context.lagda.tex
revised typing contexts

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.Product using (_×_; _,_; ∃; ∃-syntax)
open import Data.List.Base using (List; []; _∷_; [_]; _++_)
open import Relation.Unary
open import Relation.Binary.PropositionalEquality using (_≡_; refl; cong)

open import Type
\end{code}

\subsection{Context Representation}
\label{sec:context-agda}

We are going to adopt a nameless representation of channels hence typing
contexts are represented using lists of types, where the (polymorphic)
\AgdaDatatype{List} type and its constructors \AgdaInductiveConstructor{[]} and
\AgdaInductiveConstructor{\_∷\_} are defined in the module
\AgdaModule{Data.List} of Agda's standard library. We will keep using
$\ContextC$, $\ContextD$ and $\ContextE$ to range over typing contexts as we
have done in \Cref{sec:typing-rules}, even though in the Agda formalization they
are lists instead of finite maps.

\begin{code}
Context : Set
Context = List Type
\end{code}

The most important operation concerning typing contexts is \emph{splitting}. The
splitting of $\ContextC$ into $\ContextD$ and $\ContextE$, which we denote by
$\Splitting\ContextC\ContextD\ContextE$, represents the fact that $\ContextC$
contains all the types contained in $\ContextD$ and $\ContextE$, preserving both
their overall \emph{multiplicity} and also their relative \emph{order} within
$\ContextD$ and $\ContextE$. A \emph{proof} of
$\Splitting\ContextC\ContextD\ContextE$ shows how the types in $\ContextC$ are
distributed in $\ContextD$ and $\ContextE$ from left to right.

\begin{code}[hide]
infix  4 _≃_+_
infixr 8 _─∗_
infixr 9 _∗_
\end{code}
\begin{code}
data _≃_+_ : Context → Context → Context → Set where
  •   : [] ≃ [] + []
  <_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ A ∷ Δ + Θ
  >_  : ∀{A Γ Δ Θ} → Γ ≃ Δ + Θ → A ∷ Γ ≃ Δ + A ∷ Θ
\end{code}

When splitting a context $\Context$ into $\ContextD + \ContextE$, for each type
in $\Context$ we use one of the prefix operators \AgdaInductiveConstructor{<}
and \AgdaInductiveConstructor{>} to indicate whether the type is meant to be
placed in $\ContextD$ or $\ContextE$.
%
For example, below is a proof of the splitting $\Splitting{[A, B, C,
D]}{[B]}{[A, C, D]}$.

\begin{code}[hide]
module _ where
  postulate A B C D : Type
\end{code}
\begin{code}
  splitting-example₁ : (A ∷ B ∷ C ∷ D ∷ []) ≃ (B ∷ []) + (A ∷ C ∷ D ∷ [])
  splitting-example₁ = > < > > •
\end{code}

Context splitting enjoys a number of expected properties. In particular, it is
easy to see that splitting is commutative and that the empty context/list is
both a left and right unit of splitting.

\begin{code}
+-comm    : ∀{Γ Δ Θ} → Γ ≃ Δ + Θ → Γ ≃ Θ + Δ
≫         : ∀{Γ} → Γ ≃ [] + Γ
≪         : ∀{Γ} → Γ ≃ Γ + []
\end{code}
\begin{code}[hide]
+-comm • = •
+-comm (< p) = > (+-comm p)
+-comm (> p) = < (+-comm p)

++≃+ : ∀{Γ Δ} → Γ ++ Δ ≃ Γ + Δ
++≃+ {[]}    {[]}    = •
++≃+ {[]}    {_ ∷ _} = > ++≃+
++≃+ {_ ∷ _} {_}     = < ++≃+

≫ = ++≃+ {[]}

≪ = +-comm ≫
\end{code}

Context splitting is also associative in a sense that is made precise below. If
we write $\ContextD + \ContextE$ for some $\ContextC$ such that $Γ ≃ Δ + Θ$, then
we can prove that $Γ_1 + (Γ_2 + Γ_3) = (Γ_1 + Γ_2) + Γ_3$.

\begin{code}
+-assoc-r  : ∀{Γ Δ Θ Δ′ Θ′} → Γ ≃ Δ + Θ → Θ ≃ Δ′ + Θ′ →
             ∃[ Γ′ ] Γ′ ≃ Δ + Δ′ × Γ ≃ Γ′ + Θ′
+-assoc-l  : ∀{Γ Δ Θ Δ′ Θ′} → Γ ≃ Δ + Θ → Δ ≃ Δ′ + Θ′ →
             ∃[ Γ′ ] Γ′ ≃ Θ′ + Θ × Γ ≃ Δ′ + Γ′
\end{code}
\begin{code}[hide]
+-assoc-r • • = [] , • , •
+-assoc-r (< p) q with +-assoc-r p q
... | _ , p′ , q′ = _ , < p′ , < q′
+-assoc-r (> p) (< q) with +-assoc-r p q
... | _ , p′ , q′ = _ , > p′ , < q′
+-assoc-r (> p) (> q) with +-assoc-r p q
... | _ , p′ , q′ = _ , p′ , > q′

+-assoc-l p q with +-assoc-r (+-comm p) (+-comm q)
... | Δ , r , p′ = Δ , +-comm r , +-comm p′
\end{code}

% It is often the case that the context $\ContextD$ in a splitting
% $\Splitting\ContextC\ContextD\ContextE$ is a singleton list $[A]$. We use this
% particular splitting to describe the presence of a channel of type $A$ in
% $\ContextC$ and giving a name $\ContextE$ to the residual context. For this
% particular case, we introduce the dedicated notation
% $\SimpleSplitting\ContextC{A}\ContextE$ which we find clearer.

% \begin{code}
% _∋_⊳_ : Context → Type → Context → Set
% Γ ∋ A ⊳ Δ = Γ ≃ [ A ] + Δ
% \end{code}
\begin{code}[hide]
+-empty-l : ∀{Γ Δ} → Γ ≃ [] + Δ → Γ ≡ Δ
+-empty-l • = refl
+-empty-l (> p) = cong (_ ∷_) (+-empty-l p)
\end{code}

Note that, when specifying a splitting $\Splitting\ContextC{[A]}\ContextE$, it
may be convenient to use \AgdaFunction{≫} as a shortcut for a sequence of
applications of \AgdaInductiveConstructor{>} after the $A$ type has been
reached. For instance, \AgdaFunction{splitting-example₁} can be written
equivalently and in a more compact way as shown below. More usages of
\AgdaFunction{≫} will be provided in \Cref{sec:examples}.

\begin{code}[hide]
module _ where
\end{code}
\begin{code}
  splitting-example₂ : (A ∷ B ∷ C ∷ D ∷ []) ≃ [ B ] + (A ∷ C ∷ D ∷ [])
  splitting-example₂ = > < ≫
\end{code}

\begin{table}[t]
  \begin{tabular}{|lll|}
    \hline
    \textbf{Notation} & \textbf{Definition} & \textbf{Meaning} \\
    \hline
    $\AgdaFunction{∀[}~P~\AgdaFunction]$ & $\forall\{x\} \to P~x$ & implicit universality \\
    $P~\AgdaFunction{⇒}~Q$ & $\lambda x \to P~x \to Q~x$ & implication \\
    $P~\AgdaFunction{∪}~Q$ & $\lambda x \to P~x~\AgdaDatatype{⊎}~Q~x$ & disjunction \\
    $P~\AgdaFunction{∩}~Q$ & $\lambda x \to P~x~\AgdaDatatype{×}~Q~x$ & conjunction \\
    $f~\AgdaFunction{⊢}~P$ & $\lambda x \to P~(f~x)$ & update \\
    $\AgdaFunction{U}$ & $\lambda x \to \AgdaModule{Data.Unit}.\AgdaDatatype{⊤}$ & universal
    set \\
    $\AgdaFunction{⋂[}~X~\AgdaFunction{:}~A~\AgdaFunction]~P$ & $\lambda x \to (X : A) \to P~X~x$ & infinitary conjunction \\
    \hline
  \end{tabular}
  \caption{\label{tab:predicates} Useful definitions from Agda's \AgdaModule{Relation.Unary} module.}
\end{table}

Hereafter we will make extensive use of predicates over contexts. For this
reason, it is worth recalling a number of definitions from the
\AgdaModule{Relation.Unary} module in Agda's standard library in
\Cref{tab:predicates}.

Context splitting is going to be pervasive in the subsequent development, to the
point that it contributes substantially to the complexity of the formalization.
For this reason, we invest some effort in defining a few abstractions that allow
us to hide context splittings and enhance code readability.
%
Inspired by separation logic~\cite{OHearnPym99} and
following~\citet{RouvoetPoulsenKrebbersVisser20}, we define the \emph{separating
conjunction} $P \sand Q$ of two predicates $P$ and $Q$ over contexts:
\begin{code}
data _∗_ (P Q : Pred Context _) (Γ : Context) : Set where
  _⟨_⟩_ : ∀{Δ Θ} → P Δ → Γ ≃ Δ + Θ → Q Θ → (P ∗ Q) Γ
\end{code}

The predicate $P \sand Q$ holds for those contexts $\ContextC$ that can be split
into $\ContextD$ and $\ContextE$ where $P$ holds for $\ContextD$ and $Q$ holds
for $\ContextE$.
%
The constructor \AgdaInductiveConstructor{\_⟨\_⟩\_} has three explicit arguments
witnessing the splitting $\Splitting\ContextC\ContextD\ContextE$ along with
$P~\ContextD$ and $Q~\ContextE$.
%
The use of the metavariables $P$ and $Q$ for denoting predicates over contexts
is not without reason. As we will see shortly, in our formalization processes
are indeed predicates over typing contexts, albeit not the only ones.

Along with \AgdaDatatype{∗} we define the \emph{separating implication} (also
known as ``magic wand'')
\begin{code}
_─∗_ : Pred Context _ → Pred Context _ → Context → Set
(P ─∗ Q) Δ = ∀{Θ Γ} → Γ ≃ Δ + Θ → P Θ → Q Γ
\end{code}
and prove that \AgdaFunction{─∗} can be used to curry \AgdaDatatype{∗}:
\begin{code}
curry∗ : ∀{P Q R} → ∀[ P ∗ Q ⇒ R ] → ∀[ P ⇒ Q ─∗ R ]
curry∗ F px σ qx = F (px ⟨ σ ⟩ qx)
\end{code}

To conclude the modeling of typing contexts, we define a predicate
\AgdaDatatype{Un} that holds for \emph{unrestricted} contexts, namely those
contexts solely made of types of the form $?A$. We will need this predicate in
the definition of a server, which must obey the typing rule \ServerRule of
\Cref{tab:typing-rules}.

\begin{code}
data Un : Context → Set where
  un-[]  : Un []
  un-∷   : ∀{A} → ∀[ Un ⇒ (`? A ∷_) ⊢ Un ]
\end{code}

It is easy to prove that $\Splitting\ContextC\ContextD\ContextE$ is unrestricted
if so are $\ContextD$ and $\ContextE$:

\begin{code}
∗-un : ∀[ Un ∗ Un ⇒ Un ]
\end{code}
\begin{code}[hide]
∗-un (un-[] ⟨ • ⟩ un-[])    = un-[]
∗-un (un-∷ un ⟨ < σ ⟩ un′)  = un-∷ (∗-un (un ⟨ σ ⟩ un′))
∗-un (un ⟨ > σ ⟩ un-∷ un′)  = un-∷ (∗-un (un ⟨ σ ⟩ un′))
\end{code}