### 2025-12-24 paper/Safety.lagda.tex
revised safety

\subsection{Type Safety}
\label{sec:safety-agda}

\begin{code}[hide]
{-# OPTIONS --rewriting #-}
open import Data.List.Base using ([]; _∷_; [_])
open import Relation.Unary

open import Type
open import Context
open import Process
open import Congruence
open import DeadlockFreedom
\end{code}

We have seen that type safety is a simple instance of deadlock freedom, which is
made even simpler to formalize in our development where processes are
intrinsically typed. We start by defining reduction contexts as processes with a
single hole. In our intrinsically-typed formalization, reduction contexts are
parameterized by the typing context $\ContextD$ of the hole, which is invariant,
and indexed by the typing context $\ContextC$ of the whole reduction context:

\begin{code}
data ReductionContext (Δ : Context) : Context → Set where
  hole   : ReductionContext Δ Δ
  cut-l  : ∀{A} → ∀[  ((A ∷_) ⊢ ReductionContext Δ) ∗ ((dual A ∷_) ⊢ Proc) ⇒
                      ReductionContext Δ ]
  cut-r  : ∀{A} → ∀[  ((A ∷_) ⊢ Proc) ∗ ((dual A ∷_) ⊢ ReductionContext Δ) ⇒
                      ReductionContext Δ ]
\end{code}

Note the presence of two constructors \AgdaInductiveConstructor{cut-l} and
\AgdaInductiveConstructor{cut-r} which distinguish on whether the hole is found
in the left (respectively, right) sub-context, as per the grammar of reduction
contexts in \Cref{sec:properties}.

Substitution inside a reduction context $\ReductionContext$ is a straightforward
function $\AgdaFunction{\_⟦\_⟧}$ that operates recursively on the structure of
$\ReductionContext$:

\begin{code}
_⟦_⟧ : ∀{Γ Δ} → ReductionContext Δ Γ → Proc Δ → Proc Γ
hole               ⟦ P ⟧ = P
cut-l (𝒞 ⟨ p ⟩ Q)  ⟦ P ⟧ = cut ((𝒞 ⟦ P ⟧) ⟨ p ⟩ Q)
cut-r (Q ⟨ p ⟩ 𝒞)  ⟦ P ⟧ = cut (Q ⟨ p ⟩ (𝒞 ⟦ P ⟧))
\end{code}

A process $P$ is well formed if every unguarded sub-process $Q$ in it is alive.

\begin{code}
WellFormed : ∀{Γ} → Proc Γ → Set
WellFormed {Γ} P = ∀{Δ} {𝒞 : ReductionContext Δ Γ} {Q : Proc Δ} →
                   P ⊒ (𝒞 ⟦ Q ⟧) → Alive Q
\end{code}

The proof of type safety ends up being a trivial application of
\AgdaFunction{deadlock-freedom}, since the process $Q$ in the hole of a
reduction context is well typed by construction.

\begin{code}
type-safety : ∀{Γ} (P : Proc Γ) → WellFormed P
type-safety P {_} {_} {Q} _ = deadlock-freedom Q
\end{code}