### 2025-12-24 paper/abstract.tex
revised intro, british spelling

\abstract{%
    We describe an Agda formalisation of a process calculus based on classical
    linear logic that supports the modeling of binary sessions. The
    formalisation turns out to be remarkably compact despite the use of
    ``plain'' context splitting for the handling of linear resources. We argue
    that the logical nature of the calculus and the use of explicit
    continuations contribute to the simplicity of its formalisation.
    %
}