### 2025-12-25 paper/abstract.tex
revised abstract and intro

\abstract{%
    We describe the Agda formalisation of a process calculus based on classical
    linear logic that supports the modeling of binary sessions through their
    encoding with explicit continuation channels. The formalisation turns out to
    be remarkably compact despite the use of context splitting for the proper
    management of linear channels. We argue that the logical nature of the
    calculus and the use of explicit continuations are contributing factors to
    the simplicity of its formalisation.
    %
}