### 2025-12-26 paper/agda.tex

\section{Agda Formalisation}
\label{sec:agda}

In this section we describe the formalisation of \LCC in Agda. We present in
detail only some key parts of the code including the representation of types and
processes, the definition of the operational semantics and the statement of the
main results. The complete source code is available in \LCC's public repository.

\input{Type}
\input{Context}
\input{Permutations}
\input{Process}
\input{Congruence}
\input{Reduction}
\input{DeadlockFreedom}
\input{Safety}
\input{Examples}