### 2025-12-27 paper/abstract.tex
tuning

\abstract{%
    The formalisation of session calculi is considered challenging because of
    the management of channels, which are linear resources that cannot be
    discarded or duplicated and whose type changes over time, as input/output
    operations are performed on them.
    %
    Context splitting, the ``naive'' channel management technique directly
    related to the way session type theories are usually written with pen and
    paper, is often considered a hindrance and a notable source of complexity,
    to the point that several alternative approaches have been recently
    proposed.
    %
    In this paper we describe the Agda formalisation of a process calculus based
    on classical linear logic that supports the modeling of binary sessions
    through their encoding with explicit continuation channels. The
    formalisation turns out to be remarkably compact despite the adoption of
    context splitting. We argue that the logical nature of the calculus and the
    use of explicit continuations are contributing factors to the simplicity of
    its formalisation.
    %
}