### 2025-12-29 paper/agda.tex

\section{Agda Formalisation}
\label{sec:agda}

In this section we describe the formalisation of \LCC in Agda. Each of the
following sub-sections matches one of the modules of the formalisation. We
present in detail only some key parts of the code including the representation
of types and processes, the definition of the operational semantics and the
statement of the main results. The complete source code is available in \LCC's
public repository~\cite{PadovaniRaffaelli25}.

\input{Type}
\input{Context}
\input{Permutations}
\input{Process}
\input{Congruence}
\input{Reduction}
\input{DeadlockFreedom}
\input{Safety}
\input{Examples}